\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-05 R/classify_species.Rmd
working on confusion matrix figures

rf_train_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True species for Training Set") + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none")

# save confusion matrix to file
ggsave(filename = here::here("analysis","figures", "confMat_species.pdf"), width = 14, height = 14)
```

Use the randomForest model to predict validation set species. Visualize classification results across species for the VALIDATION SET using ggplot:

@@

rf_valid_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True SPECIES for Validation Set") + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove legend entirely
  theme(legend.position = "none")

@@

```{r}
rf_classifier_genus <- readRDS("output/rf_classifier_genus_mtry27_ntree100.rds")
```

@@

rf_valid_results_genus %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos_genus) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True GENUS for Validation Set") + 
  scale_fill_gradient(low = "white", high = "#fc8d59") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none")
```

>>>>

rf_train_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True species for Training Set") + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none")
```

Use the randomForest model to predict VALIDATION set SPECIES. Visualize classification results across species for the VALIDATION SET using ggplot:

@@

rf_valid_results %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", 
       y = "Predicted"
       #, title = "RandomForest Classification Confusion Matrix: Predicted vs. True SPECIES for Validation Set"
       ) + 
  scale_fill_gradient(low = "white", high = "#43a2ca") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove legend entirely
  theme(legend.position = "none",
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

ggsave(filename = here::here("analysis","figures", "confMat_species_valid.pdf"), width = 15, height = 15)
```

@@

```{r}
rf_classifier_genus <- readRDS(here::here("R","output","rf_classifier_genus_mtry27_ntree100.rds"))
```

@@

rf_valid_results_genus %>% 
  # count number of each obs/pred combination
  count(obs,pred) %>% 
  # fill in combinations with counts of zero
  complete(obs, pred, fill = list(n = 0)) %>% 
  # add rows witth species combinations not in predicted set
  right_join(all_combos_genus) %>% 
  # set NA values to 0 in the n column
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  # set raster cell fill color based on count
  ggplot(aes(x = obs, y = fct_rev(pred), fill = n)) + 
  geom_raster() + 
  # add boxes around diagonal cells
  geom_tile(aes(color = obs == pred, width = 0.95, height = 0.95), size = 0.5) + 
  scale_color_manual(values = c(NA, "black")) + 
  # rotate x axis text 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Truth", y = "Predicted", title = "RandomForest Classification Confusion Matrix: Predicted vs. True GENUS for Validation Set") + 
  scale_fill_gradient(low = "white", high = "#fc8d59") + 
  # add count in each cell
  geom_text(aes(label = n, alpha = n > 0)) + 
  # remove the color legend
  #guides(color = "none")
  # remove legend entirely
  theme(legend.position = "none")

# save figure to file 
ggsave(filename = here::here("analysis","figures", "confMat_genus_valid.pdf"), width = 10, height = 10)
```

### 2020-11-05 analysis/paper/paper.Rmd
working on confusion matrix figures

Figure - confusion matrices
Figure - variable importance vs wavelength
Figure - multi-year observations of individual plants?

@@

>>>>

Confusion matrix for species predictions (Figure \@ref(fig:fig-confMatSpecies))
Confusion matrix for genus predictions (Figure \@ref(fig:fig-confMatGenus))

Figure - variable importance vs wavelength
Figure - multi-year observations of individual plants?

@@

## Figure 4 {-}

```{r fig-confMatSpecies, fig.cap ="Confusion matrix Species", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "confMat_species_valid.pdf"))
```

## Figure 5 {-}

```{r fig-confMatGenus, fig.cap ="Confusion matrix Species", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "confMat_genus_valid.pdf"))
```
